### 2021-06-07 appendices/explore.Rmd

---
output: pdf_document
---

```{r setup, cache = FALSE, message=FALSE, include=FALSE}
# R dependencies
library(tidyverse)
library(patchwork)
library(reticulate)

## Python dependencies loaded via R
sb3      = import("stable_baselines3")
torch    = import("torch")
gym      = import("gym")
gym_fishing = import("gym_fishing")
gym_conservation = import("gym_conservation")

source("../R/plotting-helpers.R")
ggplot2::theme_set(ggplot2::theme_bw())
scale_colour_discrete = function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete = function(...) ggthemes::scale_fill_solarized()

knitr::opts_chunk$set(echo=FALSE)
```

```{r}

sims_all <- function(cache, env) {

  names(cache) <- gsub("_\\w*tuned\\.zip", "", basename(cache))
  algos <- stringr::str_to_upper(names(cache))

  names(algos) <- algos
  sims_df <- map_dfr(algos, function(algo){
    i <- which(algo==algos)
    agent <- sb3[[algo]]$load(cache[[i]])
    env$simulate(agent, 100L)
  }, .id = "model")

  opt <- gym_fishing$models$escapement(env)
  opt_sims <- env$simulate(opt, reps = 100L)%>% mutate(model="optimal")
  bind_rows(opt_sims, sims_df) 
}

policyfns <- function(cache, env) {

  names(cache) <- gsub("_\\w*tuned\\.zip", "", basename(cache))
  algos <- stringr::str_to_upper(names(cache))
  names(algos) <- algos
  policy_df <- map_dfr(algos, function(algo){
    i <- which(algo==algos)
    agent <- sb3[[algo]]$load(cache[[i]])
    env$policyfn(agent, 1L)
  }, .id = "model")

  opt <- gym_fishing$models$escapement(env)
  opt_policy <- env$policyfn(opt)%>% mutate(model="optimal")
  bind_rows(opt_policy, policy_df) 
}

#  policy_df = env$policyfn(agent) %>% mutate(model="RL") %>% bind_rows(opt_policy)
```

# Deterministic Environment, Untuned agents

```{r}
env = gym$make("fishing-v1", sigma = 0)
untuned_det <- fs::dir_ls("../python/fishing/determinstic/cache/", regexp = ".*_untuned\\.zip")

df <- sims_all(untuned_det, env)
pf <- policyfns(untuned_det, env)
plot_sims(df) / (plot_policy(pf) + plot_reward(df))
```

# Stochastic Environment, Untuned Agents

```{r}
env = gym$make("fishing-v1", sigma = 0.1)
cache <- fs::dir_ls("../python/fishing/stochastic/cache/", regexp = ".*_untuned\\.zip")

df <- sims_all(cache, env)
pf <- policyfns(cache, env)
plot_sims(df) / (plot_policy(pf) + plot_reward(df))
```

# Deterministic Environment, Tuned Agents

```{r}
env = gym$make("fishing-v1", sigma = 0)
tuned_det <- fs::dir_ls("../python/fishing/determinstic/cache/", regexp = ".*_tuned\\.zip")

df <- sims_all(tuned_det, env)
pf <- policyfns(tuned_det, env)
plot_sims(df) / (plot_policy(pf) + plot_reward(df))
```

# Stochastic Environment, Tuned Agents

```{r}
env = gym$make("fishing-v1", sigma = 0.1)
tuned_stoch <- fs::dir_ls("../python/fishing/stochastic/cache/", regexp = ".*_tuned\\.zip")

df <- sims_all(tuned_stoch, env)
pf <- policyfns(tuned_stoch, env)
plot_sims(df) / (plot_policy(pf) + plot_reward(df))
```