### 2021-06-07 appendices/appendix-D-data.Rmd
updated examples

---
title: "Appendix D: Accessing Empirical Data for Management Comparison"
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Melissa Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Kari Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu

bibliography: ../manuscript/references.bib

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}

output: rticles::arxiv_article
---

This appendix provides reproducible code used to access and clean data from the R.A. Myers Legacy Stock Assessment Database used in for empirical comparison in the hindcasts shown in the main text.

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(8)
txtcolor <- "#586e75"
```

```{r setup, cache = FALSE, message=FALSE}
# R dependencies
library(tidyverse)
library(patchwork)
library(reticulate)

## Python dependencies loaded via R
sb3         <- import ("stable_baselines3")
gym         <- import ("gym")
gym_fishing <- import("gym_fishing")
```

Access data from the RAM Legacy Stock Assessment Database. (2020). Version v4.491 [Data set]. Zenodo. http://doi.org/10.5281/zenodo.3676088.
Note that DOIs resolve only to HTML landing pages.  
For programmatic access to the data, we resolve content by it's content hash instead:

```{r}
library(contentid)
ramzip <- resolve("hash://md5/fac27d8b7876df29afaeadcd7f5a3b77")
unzip(ramzip, exdir = "ramlegacy")
load("ramlegacy/RAMLDB v4.491/DB Files With Assessment Data/R Data/DBdata[asmt][v4.491].RData")
```

Relevant data is stored across several tables, following relational database practices.  
Before analysis we join these tables by shared primary key:

```{r}
ramlegacy <- 
  timeseries_values_views %>%
  select(stockid, stocklong, year, SSB, TC) %>%
  left_join(stock) %>%
  left_join(area) %>%
  select(stockid, scientificname, 
         commonname, areaname, country, year, 
         SSB, TC) %>%
  left_join(timeseries_units_views %>%
              rename(TC_units = TC, SSB_units = SSB)) %>%
  select(scientificname, commonname, 
         stockid, areaname, country, year, 
         SSB, TC, SSB_units, TC_units)
```

Let's filter out missing data, non-matching units, and obvious reporting errors (catch exceeding total spawning biomass), then we re-scale each series into the 0,1 by appropriate choice of units.
We store the processed data locally for future use.

```{r}
stock_ids <- c("ARGHAKENARG")

fish <- ramlegacy %>% 
  filter(stockid %in% stock_ids) %>% 
  filter(!is.na(SSB), !is.na(TC)) %>%
  filter(SSB_units == "MT", TC_units=="MT") %>% 
  filter(SSB > TC) %>%
  select(-SSB_units, -TC_units) %>% 
  group_by(stockid) %>%
  mutate(scaled_catch = TC / max(SSB),
         scaled_biomass = SSB / max(SSB))

hake <- fish  %>% filter(commonname=="Argentine hake")
write_csv(hake, "../data/hake.csv")
```