### 2022-01-27 appendices/failure_plot.Rmd
Stashing here for now

---
title: "failure_plot"
output: html_document
---

```{r message=FALSE}
#install.packages("renv")
renv::restore()
```

```{r setup, cache = FALSE, message=FALSE}
# R dependencies
library(tidyverse)
library(patchwork)
library(reticulate)

## Python dependencies loaded via R
sb3      = import("stable_baselines3")
torch    = import("torch")
gym      = import("gym")
gym_fishing = import("gym_fishing")
gym_conservation = import("gym_conservation")

source("../R/plotting-helpers.R")
```

```{r knitr, include = FALSE}
# This chunk controls plot aesthetics only, and can be omitted with no material change
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())

scale_colour_discrete = function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete = function(...) ggthemes::scale_fill_solarized()
pal = ggthemes::solarized_pal()(8)
txtcolor = "#586e75"

colors <- set_names(c(pal[c(1,3,8,2)], txtcolor, pal[6]),
                    c("Optimal", "TD3_untuned", "TD3_tuned", "RL", "historical", "steady-state"))
scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
```

```{r}
## reproducible settings
np = import("numpy")
seed = 24L # integer
np$random$seed(seed)

set.seed(seed)
```

```{r }
## initialize the environment
env_high = gym$make("fishing-v1", sigma=0.1, init_state=1.5)
env_low = gym$make("fishing-v1", sigma=0.1)
```

```{r}
# Simulate under the optimal solution (given the model)
opt = gym_fishing$models$escapement(env_low)
opt_sims_high = env_high$simulate(opt, reps = 100L)
opt_sims_low = env_low$simulate(opt, reps = 100L)
opt_policy = env_low$policyfn(opt)
```

```{r}
# add a column indicating the model
sims_df_high <- opt_sims_high %>% mutate(model = "Optimal")
sims_df_low <- opt_sims_low %>% mutate(model = "Optimal")
policy_df <- opt_policy  %>% mutate(model = "Optimal")
```

```{r}
agent = sb3$TD3$load("../python/cache/td3_untuned")
agent_sims_high = env_high$simulate(agent, reps = 100L)
agent_sims_low = env_low$simulate(agent, reps = 100L)
agent_policy = env_low$policyfn(agent, reps = 5L)
```

```{r}
# stack result simulations with those above and plot:
sims_df_high <- agent_sims_high %>% 
  mutate(model = "TD3_untuned") %>% 
  bind_rows(sims_df_high)

# stack result simulations with those above and plot:
sims_df_low <- agent_sims_low %>% 
  mutate(model = "TD3_untuned") %>% 
  bind_rows(sims_df_low)

policy_df <- agent_policy  %>% 
  mutate(model = "TD3_untuned")  %>% 
  bind_rows(policy_df)
```

```{r}
td3 <- sb3$TD3$load("../python/cache/td3_tuned.zip")
td3_sims_high = env_high$simulate(td3, reps = 100L)
td3_sims_low = env_low$simulate(td3, reps = 100L)
td3_policy = env_low$policyfn(td3)
```

```{r}
# stack result simulations with those above and plot:
sims_df_high <- td3_sims_high %>% 
  mutate(model = "TD3_tuned") %>% 
  bind_rows(sims_df_high)
sims_df_low <- td3_sims_low %>% 
  mutate(model = "TD3_tuned") %>% 
  bind_rows(sims_df_low)

policy_df <- td3_policy  %>% 
  mutate(model = "TD3_tuned")  %>% 
  bind_rows(policy_df)

( plot_sims(sims_df_low, "none") + plot_sims(sims_df_high, "right")) / ( plot_policy(policy_df))
```